Unknown

### 2024-04-01 ms/ms.qmd
working on draft of data cleaning methods

---
title: Unknown
format:
  pdf:
    keep-tex: true  
    toc: false
header-includes: |
  \addtokomafont{disposition}{\rmfamily}
  \usepackage{siunitx}
  \newcommand{\aax}{$\mathrm{AA}$}
  \newcommand{\Aamphi}{$A_{\mathrm{amphi}}$}
  \newcommand{\Ahypo}{$A_{\mathrm{hypo}}$}
  \newcommand{\agcurve}{$A \textendash g_\text{sw}$}
  \newcommand{\caequals}[1]{$C_\mathrm{a} = \qty{#1}{\micro\mol\raiseto{-1}\mol}$}
  \newcommand{\gsw}{$g_\text{sw}$}
  \newcommand{\ppfd}{$\mathrm{PPFD}$}
  \newcommand{\ppfdequals}[1]{$\mathrm{PPFD} = \qty{#1}{\micro\mol\raiseto{-2}\meter\raiseto{-1}\second}$}
  \newcommand{\rh}{$\mathrm{RH}$}
  \newcommand{\rhequals}[1]{$\mathrm{RH} = #1\%$}
  \newcommand{\tleafequals}[1]{$T_\mathrm{leaf} = \qty{#1}{\degreeCelsius}$}
mainfont: Times New Roman
bibliography: solanum-aa.bib
csl: https://www.zotero.org/styles/science
---

@@

@@

@@

- models fit in Stan/brms
- warmup and sampling iterations were chosen so that parameter ESS > 1000 and  $\hat{R} < 1.01$ [@gelman_inference_1992] and the effective sample size (ESS) was greater than $10^3$.

>>>>

---
title: Unknown
format:
  pdf:
    keep-tex: true  
    toc: false
header-includes: |
  \addtokomafont{disposition}{\rmfamily}
  \usepackage{siunitx}
  \newcommand{\aax}{$\mathrm{AA}$}
  \newcommand{\Aamphi}{$A_{\mathrm{amphi}}$}
  \newcommand{\Ahypo}{$A_{\mathrm{hypo}}$}
  \newcommand{\agcurve}{$A \textendash g_\text{sw}$}
  \newcommand{\ca}{$C_\mathrm{a}$}
  \newcommand{\caequals}[1]{$C_\mathrm{a} = \qty{#1}{\micro\mol\raiseto{-1}\mol}$}
  \newcommand{\cabetween}[2]{#1 < $C_\mathrm{a} < \qty{#2}{\micro\mol\raiseto{-1}\mol}$}
  \newcommand{\gsw}{$g_\text{sw}$}
  \newcommand{\ppfd}{$\mathrm{PPFD}$}
  \newcommand{\ppfdequals}[1]{$\mathrm{PPFD} = \qty{#1}{\micro\mol\raiseto{-2}\meter\raiseto{-1}\second}$}
  \newcommand{\rh}{$\mathrm{RH}$}
  \newcommand{\rhequals}[1]{$\mathrm{RH} = #1\%$}
  \newcommand{\tleafequals}[1]{$T_\mathrm{leaf} = \qty{#1}{\degreeCelsius}$}
mainfont: Times New Roman
bibliography: solanum-aa.bib
csl: https://www.zotero.org/styles/science
---

@@

### Cleaning RH curves
n_rh_curve_raw = 2400 # export from adaptive-amphistomy/r/07_process-licor.R
n_points_per_rh_curves_raw = 100 # export from adaptive-amphistomy/r/07_process-licor.R
ca_min = 410 # export from adaptive-amphistomy/r/10_filter-rh-curves.R
ca_max = 420 # export from adaptive-amphistomy/r/10_filter-rh-curves.R
n_min_points_per_rh_curve = 10 # export from adaptive-amphistomy/r/10_filter-rh-curves.R

@@

### Cleaning \agcurve{} curves

The raw data set consisted of `r n_rh_curve_raw` \agcurve{} curves with an average of `r n_points_per_rh_curves_raw`. Principled, consistent, manual curation of a data set this size is not feasible. Therefore, we automated data cleaning using custom *R* scripts. Cleaning is divided into XXX sequential steps:

1. remove unreliable data points
2. remove hysteretic portion of \agcurve{} curves at low \gsw{}
3. remove outliers within each \agcurve{} curve
4. remove replicates with no overlap between amphi and pseudohypo \agcurve{} curves
5. remove replicates with outlier \aax values

#### Remove unreliable data points

*Rationale*: Unreliable data points consisted of those where chamber [CO$_2$] was unstable and therefore measurements are not biologically meaningful.

*Procedure*: We retained data points where \cabetween{`r ca_min`}{`r ca_max`}.

#### Remove hysteretic portion of \agcurve{} curves at low \gsw{}

*Rationale*: In most \agcurve{} curves, we observed a hysteretic response at low \gsw. After \gsw{} and $A$ declined simultaneously, $A$ increased slightly as \gsw{} continued to decline or stabilize, indicating some leaf acclimation to low \rh. We removed this portion of the curve to focus curve-fitting on the primary domain where $A$ increases monotonically with \gsw{}.

*Procedure*: For each curve, we removed data points after \gsw{} had reached its minimum unless there were fewer than `r n_min_points_per_rh_curve` data points remaining.

WORKING HERE

@@

- models fit in Stan/brms
- we chose the number of chains, warmup and sampling iterations, and maximum treedepth so that parameter estimates converged ($\hat{R} < 1.01$ [@gelman_inference_1992]) and the effective sample size (ESS) for each parameter was greater than $10^3$.

### 2024-04-01 ms/ms.tex
working on draft of data cleaning methods

\KOMAoption{captions}{tableheading}
\addtokomafont{disposition}{\rmfamily}
\usepackage{siunitx}
\newcommand{\aax}{$\mathrm{AA}$}
\newcommand{\Aamphi}{$A_{\mathrm{amphi}}$}
\newcommand{\Ahypo}{$A_{\mathrm{hypo}}$}
\newcommand{\agcurve}{$A \textendash g_\text{sw}$}
\newcommand{\caequals}[1]{$C_\mathrm{a} = \qty{#1}{\micro\mol\raiseto{-1}\mol}$}
\newcommand{\gsw}{$g_\text{sw}$}
\newcommand{\ppfd}{$\mathrm{PPFD}$}
\newcommand{\ppfdequals}[1]{$\mathrm{PPFD} = \qty{#1}{\micro\mol\raiseto{-2}\meter\raiseto{-1}\second}$}
\newcommand{\rh}{$\mathrm{RH}$}
\newcommand{\rhequals}[1]{$\mathrm{RH} = #1\%$}
\newcommand{\tleafequals}[1]{$T_\mathrm{leaf} = \qty{#1}{\degreeCelsius}$}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

@@

@@

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1972}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3099}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2535}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2394}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
hypothesis
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
native light habitat
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
light treatment
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
light intensity
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
acclimatory & cor(PAR, AA) = 0 & AAsun = AAshade & AA2000 \textgreater{}
AA150 \\
plastic & cor(PAR, AAsun) \textgreater{} 0 & AAsun \textgreater{}
AAshade & AA2000 = AA150 \\
constitutive & cor(PAR, AA) \textgreater{} 0 & AAsun = AAshade & AA2000
= AA150 \\
\end{longtable}

@@

\paragraph{Internal leaf anatomy}\label{internal-leaf-anatomy}

Add information on internal leaf anatomy, if using

@@

@@

>>>>

\KOMAoption{captions}{tableheading}
\addtokomafont{disposition}{\rmfamily}
\usepackage{siunitx}
\newcommand{\aax}{$\mathrm{AA}$}
\newcommand{\Aamphi}{$A_{\mathrm{amphi}}$}
\newcommand{\Ahypo}{$A_{\mathrm{hypo}}$}
\newcommand{\agcurve}{$A \textendash g_\text{sw}$}
\newcommand{\ca}{$C_\mathrm{a}$}
\newcommand{\caequals}[1]{$C_\mathrm{a} = \qty{#1}{\micro\mol\raiseto{-1}\mol}$}
\newcommand{\cabetween}[2]{#1 < $C_\mathrm{a} < \qty{#2}{\micro\mol\raiseto{-1}\mol}$}
\newcommand{\gsw}{$g_\text{sw}$}
\newcommand{\ppfd}{$\mathrm{PPFD}$}
\newcommand{\ppfdequals}[1]{$\mathrm{PPFD} = \qty{#1}{\micro\mol\raiseto{-2}\meter\raiseto{-1}\second}$}
\newcommand{\rh}{$\mathrm{RH}$}
\newcommand{\rhequals}[1]{$\mathrm{RH} = #1\%$}
\newcommand{\tleafequals}[1]{$T_\mathrm{leaf} = \qty{#1}{\degreeCelsius}$}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

@@

\subsection{Authors}\label{authors}

definite: Wei Shen Lim, Dachuan Wang, Chris Muir question: Sam McKlin,
Tom Buckley

@@

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1972}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3099}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2535}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2394}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
hypothesis
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
native light habitat
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
light treatment
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
light intensity
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
acclimatory & cor(PAR, AA) = 0 & AAsun = AAshade & AA2000 \textgreater{}
AA150 \\
plastic & cor(PAR, AA) = 0 & AAsun \textgreater{} AAshade & AA2000 =
AA150 \\
constitutive & cor(PAR, AA) \textgreater{} 0 & AAsun = AAshade & AA2000
= AA150 \\
\end{longtable}

@@

@@

\subsubsection{\texorpdfstring{Cleaning \agcurve{}
curves}{Cleaning  curves}}\label{cleaning-curves}

The raw data set consisted of 2400 \agcurve{} curves with an average of
100. Principled, consistent, manual curation of a data set this size is
not feasible. Therefore, we automated data cleaning using custom
\emph{R} scripts. Cleaning is divided into XXX sequential steps:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  remove unreliable data points
\item
  remove hysteretic portion of \agcurve{} curves at low \gsw{}
\item
  remove outliers within each \agcurve{} curve
\item
  remove replicates with no overlap between amphi and pseudohypo
  \agcurve{} curves
\item
  remove replicates with outlier \aax values
\end{enumerate}

\paragraph{Remove unreliable data
points}\label{remove-unreliable-data-points}

\emph{Rationale}: Unreliable data points consisted of those where
chamber {[}CO\(_2\){]} was unstable and therefore measurements are not
biologically meaningful.

\emph{Procedure}: We retained data points where \cabetween{410}{420}.

\paragraph{\texorpdfstring{Remove hysteretic portion of \agcurve{}
curves at low
\gsw{}}{Remove hysteretic portion of  curves at low }}\label{remove-hysteretic-portion-of-curves-at-low}

\emph{Rationale}: In most \agcurve{} curves, we observed a hysteretic
response at low \gsw. After \gsw{} and \(A\) declined simultaneously,
\(A\) increased slightly as \gsw{} continued to decline or stabilize,
indicating some leaf acclimation to low \rh. We removed this portion of
the curve to focus curve-fitting on the primary domain where \(A\)
increases monotonically with \gsw{}.

\emph{Procedure}: For each

\subsubsection{Data analysis}\label{data-analysis}

\paragraph{General points (not sure where this goes
yet)}\label{general-points-not-sure-where-this-goes-yet}

\begin{itemize}
\tightlist
\item
  models fit in Stan/brms
\item
  we chose the number of chains, warmup and sampling iterations, and
  maximum treedepth so that parameter estimates converged
  (\(\hat{R} < 1.01\) (\emph{8})) and the effective sample size (ESS)
  for each parameter was greater than \(10^3\).
\end{itemize}

We fit Bayesian models using the \emph{R} package \textbf{brms} version
2.20.4 (\emph{9}) with MCMC sampling in \emph{Stan} (\emph{10}). We used
CmdStan version 2.33.1 and \textbf{cmdstanr} version 0.7.1 (\emph{11})
to interface with \emph{R} version 4.3.2 (\emph{12}). We sampled the
posterior distribution from 4 chains with 1000 iterations each after
1000 warmup iterations per chain. We estimated parameters and confidence
intervals as the median and 95\% quantile intervals of the posterior,
respectively.

\paragraph{Parameter estimates (this is a provisional
section)}\label{parameter-estimates-this-is-a-provisional-section}

Base model (need to add indices):

\[\text{log(A)} = \beta_0 + \beta_1 \text{log}(g_\text{sw}) + \beta_2 \text{log}(g_\text{sw}) ^ 2 + \epsilon\]
Estimate \aax{} by integrating over range of \gsw{} overlap (need to add
indices for accession, replicate, light intensity):

\[\mathrm{AA} = \int_{\text{min(log(}g_\text{sw}))}^{\text{max(log(}g_\text{sw}))} \text{log}\bigg(\frac{A_\text{amphi}(x; \theta_\text{amphi})}{A_\text{hypo}(x; \theta_\text{hypo})}\bigg) dx\]

where \(\text{min(log(}g_\text{sw}))\) and
\(\text{max(log(}g_\text{sw}))\) are the minimum and maximum values of
\(\text{log(}g_\text{sw})\) where amphi and pseudohypo curves overlap.
\Aamphi{} is calculated from parameter estimates
\(\theta_\text{amphi} \in \{\beta_{0,\text{amphi}}, \beta_{1,\text{amphi}}, \beta_{2,\text{amphi}}\}\);
\Ahypo{} is calculated from parameter estimates
\(\theta_\text{hypo} \in \{\beta_{0,\text{hypo}}, \beta_{1,\text{hypo}}, \beta_{2,\text{hypo}}\}\)
for each curve.

\paragraph{Predictions}\label{predictions}

\begin{itemize}
\tightlist
\item
  Assimilatory:

  \begin{itemize}
  \tightlist
  \item
    average AA2000 - AA150
  \item
    among accession variation in AA2000 - AA150
  \end{itemize}
\item
  Plastic

  \begin{itemize}
  \tightlist
  \item
    average AAsun - AAshade
  \item
    among accession variation in AAsun - AAshade
  \end{itemize}
\item
  Constitutive

  \begin{itemize}
  \tightlist
  \item
    cor(PAR, AA) at accession level
  \end{itemize}
\end{itemize}

Reinforcing interactions (we only consider these because these are ones
which could explain why amphi leaves occur in high light habitats)

\begin{itemize}
\tightlist
\item
  Assimilatory * Plastic (reinforcing interaction)

  \begin{itemize}
  \tightlist
  \item
    AAsun - AAshade greater at 2000 than 150
  \item
    AA2000 - AA150 greater in sun leaves than shade leaves
  \item
    there may be var among accessions, but not correlated with PAR
  \end{itemize}
\item
  Assimilatory * Constitutive (reinforcing interaction)

  \begin{itemize}
  \tightlist
  \item
    cor(AA2000, PAR) \textgreater{} 0; cor(AA150, PAR) = 0
  \item
    AA2000 - AA150 varies among accessions, correlated with PAR
  \end{itemize}
\item
  Plastic * Constitutive (reinforcing interaction)

  \begin{itemize}
  \tightlist
  \item
    cor(AAsun, PAR) \textgreater{} 0; cor(AAshade, PAR) = 0
  \item
    AAsun - AAshade varies among accessions, correlated with PAR
  \end{itemize}
\item
  Assimilatory * Plastic * Constitutive (reinforcing interaction)

  \begin{itemize}
  \tightlist
  \item
    AAsun - AAshade greater at 2000 than 150
  \item
    AA2000 - AA150 greater in sun leaves than shade leaves
  \item
    AAsun - AAshade varies among accession, correlated with PAR
  \item
    AA2000 - AA150 varies among accession, correlated with PAR
  \item
    cor(AAsun2000, PAR) \textgreater{} cor(AAsun150, PAR) and
    \textgreater{} cor(AAshade2000, PAR) and \textgreater{}
    cor(AAshade150, PAR) in nonadditive way
  \end{itemize}
\end{itemize}

We tested competing predictions through a combination of model selection
and parameter estimation from the posterior of selected models. We
started with a base model that includes fixed effects of light intensity
and light treatment on \aax, a random effect of accession on \aax, and a
linear regression of PAR on accession-level \aax. We next compared the
fit of the base model to a more complex with an interaction between
light intensity and light treatment using LOOIC with the \textit{R}
package \textbf{loo} version 2.7.0 (\textbf{vehtari\_practical\_2017?}).
These comparisons allowed us to test the main predictions of each
hypothesis and for a reinforcing interactions between assimilatory and
plastic hypotheses (tab:predictions). However, these models cannot test
predictions when there is an interaction with the constitutive
hypothesis, because this hypothesis posits accession-levels variation in
responses to light intensity and/or light treatment. Therefore, we
tested for significant variation among accessions in light intensity and
light treatment by comparing LOOIC values of models with and without
random interactions between these factors and accession.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3182}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2727}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4091}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Hypothesis
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Prediction
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Model
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Assimilatory & & base model \\
Plastic & & base model \\
Constitutive & & base model \\
Assimilatory * Plastic & & + intensity \(\times\) treatment \\
Assimilatory * Constitutive & & + accession \(\times\) intensity
(random) \\
Plastic * Constitutive & & + accession \(\times\) treatment (random) \\
Assimilatory * Plastic * Constitutive & & + accession \(\times\)
intensity \(\times\) treatment (random) \\
\end{longtable}

\subsection{Acknowledgements}\label{acknowledgements}

Justin Alter, Max Gatlin, Joana Kim, Jenna Matsuyama, Brandon Najarian,
Kai Yasuda

@@

\bibitem[\citeproctext]{ref-gelman_inference_1992}
\CSLLeftMargin{8. }%
\CSLRightInline{A. Gelman, D. B. Rubin, Inference from iterative
simulation using multiple sequences. \emph{Statistical Science}
\textbf{7}, 457--472 (1992).}

\bibitem[\citeproctext]{ref-burkner_brms_2017}
\CSLLeftMargin{9. }%
\CSLRightInline{P.-C. Bürkner,
\href{https://doi.org/10.18637/jss.v080.i01}{\textbf{Brms} : {An}
\emph{r} {Package} for {Bayesian} {Multilevel} {Models} {Using}
\emph{stan}}. \emph{Journal of Statistical Software} \textbf{80}
(2017).}

\bibitem[\citeproctext]{ref-stan_development_team_stan_2023}
\CSLLeftMargin{10. }%
\CSLRightInline{Stan Development Team, \emph{Stan {Modeling} {Language}
{Users} {Guide} and {Reference} {Manual}} (2023;
\url{https://mc-stan.org}).}

\bibitem[\citeproctext]{ref-gabry_cmdstanr_2023}
\CSLLeftMargin{11. }%
\CSLRightInline{J. Gabry, R. Češnovar, \emph{Cmdstanr: {R} {Interface}
to '{CmdStan}'} (2023;
\href{https://mc-stan.org/cmdstanr,\%20https://discourse.mc-stan.org}{https://mc-stan.org/cmdstanr,
https://discourse.mc-stan.org}).}

\bibitem[\citeproctext]{ref-r_core_team_r:_2023}
\CSLLeftMargin{12. }%
\CSLRightInline{R Core Team, \emph{R: {A} {Language} and {Environment}
for {Statistical} {Computing}} (R Foundation for Statistical Computing,
Vienna, Austria, 2023; \url{http://www.R-project.org/}).}
